\phantomsection\addcontentsline{toc}{chapter}{Acknowledgements} % Don't remove me!
\chapter*{\vspace{-30pt}\centering{\Large Acknowledgements}}
\vspace{-10pt}
None of this work would have been possible without the endless mentoring of my supervisor, Prof. Tara Murphy, who has been guiding my research for the past 5 years and opened my eyes to the politics and drama of science. I would also like to thank David Kaplan for his constant advice and feedback (particularly the conversations on Slack at 1am) and Keith Bannister for encouraging me to think outside the box and pursue what may seem like crazy ideas.

My other supervisor, Prof. Richard Hunstead, passed away 6 months before the submission of this thesis. He was a mentor to me from the very start of my University career, and was the most enthusiastic lab supervisor I've ever seen. He was always excited be shown new results, dig up archival data, and share his vast wealth of knowledge and experience. Dick, I reckon I've done a pretty good job of this thesis, but it would be even better with a bit of feedback from your green pen.

During the course of this degree I also lost a close friend of mine. James, thanks for always being interested in my work. One of the saddest parts of finishing this thesis is knowing that you're not here to demand I explain every detail of it to you.

I've been lucky to share an office with a number of excellent people, but Andrew and Mohammad and Harry were there from start to finish. Thanks for all the help, the food, the arguments and the company. I also profusely apologise to everyone in the neighbouring offices who had to listen to our nonsense.

Emil -- you taught me everything I know about radio interferometry. Thanks for taking your time to answer all of my questions, troubleshoot data reduction and telling me all of the tips and trick that only a wizard like you could possibly know.

The follow-up of GW170817 was the most stressful experience of my life, so a big thankyou to Christene for riding that wave with me. The follow-up of GW190814 was less stressful, but I couldn't have done it without the help of Adam and the rest of the radio transients group.

I thank all of the ATNF staff, but particularly Robin for teaching me everything I know about the ATCA, Jamie for handling the bombardment of requests for observing time, and Phil for the many friendly chats we had in the hallways, and in the ATCA Portal.

I express the deepest gratitude to everyone outside of astronomy thatI've dragged along this long journey. To each and every one of my friends for listening to my rants, all the fun you've provided, and not getting too upset when I get flakey. To my Mum, Dad, sister, Grandma, Oma for their endless love and support, I couldn't have done it without you.